The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society

### 2023-12-03 manuscript/manuscript.qmd
Start big cleanup/consolidation of lit review/theory/expectations sections

# Closing space, differential enforcement, and INGO operational space

The 2017 ONGO Law's provision of two legal pathways for registration is an example of administrative crackdown [@Chaudhry:2022] an example of fit within the broader global phen

Why crack down?

Many different reasons for these restrictions and several purposes - INGOs pose a threat to regime stability, but they also provide governments with services and technical expertise. Civil society/CSO/NGO + host government definition - De Mattee and institutional taxonomy. My stuff on gatekeeping vs. program capture.

Contentiousness, regime stability, other reasons [see @Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016].

To situate our analysis of the impacts of China's ONGO Law in existing work on closing civic space, we propose a theoretical framework to explain

@@

@@

TODO: Andrew: This 2017 law fits with a broader global phenomenon of closing civic space... General crackdown stuff + China crackdown stuff + our addition

TODO: Andrew: I'll clean up/rewrite this section and the section headings, etc.

## De jure vs. de facto implementation of civil society regulations - Differential enforcement

TODO: Andrew - I'll clean up this section

@ChaudhryHeiss:2022 and the divergence in de jure and de facto regulations---the implementation doesn't have to match the written law, and often the written law can serve as a "weapon hanging on the wall that never fires" that serves as a warning to potentially uncooperative NGOs [@Kozenko:2015]. See also [@ToeplerZimmerFrohlich:2020; @Plantan:2022]

"We discuss the role of civil society organizations (CSOs) as agents of democratization and note the emergence of dual, at times apparently conflicting policy postures within authoritarian regimes (restriction and repression for some CSOs vs. financial support and opportunities for collaboration for others)." [@ToeplerZimmerFrohlich:2020]

INGO regulations are implemented and enforced differentially - difference in de jure and de facto. This differential enforcement is one mechanism to allow host governments to harness the aspects of INGO programming that align with their domestic policy preferences while limiting INGO behavior that could threaten regime stability.

Why pass laws in the first place and why enforcement differently?

@ChaudhryHeiss:2022 and the divergence in de jure and de facto regulations---the implementation doesn't have to match the written law, and often the written law can serve as a "weapon hanging on the wall that never fires" that serves as a warning to potentially uncooperative NGOs [@Kozenko:2015]. See also [@ToeplerZimmerFrohlich:2020; @Plantan:2022]

"We discuss the role of civil society organizations (CSOs) as agents of democratization and note the emergence of dual, at times apparently conflicting policy postures within authoritarian regimes (restriction and repression for some CSOs vs. financial support and opportunities for collaboration for others)." [@ToeplerZimmerFrohlich:2020]

Also, [@Plantan:2022] finds "that I find" evidence of selective implementation that reveals which groups are seen as threatening or beneficial."

## China's Overseas NGO law as a policy tool and its implications

*note* changed the sub-title to incorporate the policy tool lense too.

Such legal restrictions can be either direct (denial of legal access/cannot register) or covert (red-tape, requiring the approval of PSUs, or the uncertainty of having to filing TAs case by case).

# Empirical expectations and analytical approach

TODO: Andrew: I'll clean up this section

Operational space is determined by the interaction between NGO characteristics and regime preferences

```{r fig-theory, fig.width=6, fig.height=1.75}
#| echo: false
#| include: true
#| warning: false
#| message: false
#| fig-cap: "Diagram of our theoretical framework showing three possible determinants of INGO operational space: the contentiousness of the INGO's issue area, the inherent geographic scope of the INGO's operations, and the government's interest in and relationship with the INGO."
theory_nodes <- tribble(
  ~id, ~name, ~x, ~y, ~highlight,
  1,   "Issue contentiousness", 2, 1, FALSE,
  2,   "Operational space", 2, 0, TRUE,
  3,   "Local focus", 3, 1, FALSE,
  4,   "Government preferences", 1, 1, FALSE
)

Based on this framework, we can identify and test several theoretical expectations:

Based on extant literature and theoretical understanding, we argue that the operation space INGOs are granted is associated with the level of contentiousness their missions, i.e., the issue areas they work in. Since the level of contentiousness is also manifested by the level of difficulty they face while seeking lawful registration.

@@

@@

We lack sufficient theory or data to explain why this trend goes against our expectations, but we can again make reasonable speculations. As we saw previously with INGOs and local connections, organizations appear to strategically seek out the best avenues and methods for registering. [@Jia:2017a] Organizations that participated in the initial wave of registrations in 2017 likely had closer connections with local officials and were able to speed through the bureaucratic process registration more easily. As a preliminary test of this speculation, we rerun our main model with an interaction between years since registration and local connections to see how the effect of local connections changes over time. As seen in @fig-timing-local, early registrants in 2017 without local connections are more likely to have a larger geographic reach, registering in an average of 19--23 provinces, while their locally-focused counterparts are only expected to register in an average of 7--16 provinces. Accordingly, it may be that organizations with local connections are more likely to register early because of their pre-existing bureaucratic relationships, while organizations with a larger national reach take longer to undertake the registration process and thus register later. Another possible explanation is interpreting earlier timing in registration as more eager to register as discussed in the measurement section. INGOs that self-perceive as more contentious would be more nervous about the legal consequences of not being able to register and thus be more eager to seek registration earlier [@ChinaPhilanthropyReserchInstitute:2018]. Compared with their more "rest-assured" peers, INGOs seeking to register earlier are more likely to have reduced operational space. To understand the underlying mechanism with registration timing, further research is warranted.

@@

# Contentiousness and government relationships: The cases ofThe Nature Conservancy and Greenpeace

@@

What factors could have contributed to the contrasting experiences of TNC and Greenpeace under the ONGO Law? After analyzing the data, the working approach and relationship with the government emerged as pivotal differences between the two INGOs, which fits existing discussions on the differential treatment of INGOs based on their relationship with the authority [TO DO: add citations]. In a public interview with TNC’s RO in China about their registration experience, the chief representative stated TNC’s working approach in China has been “science-based”, using “non-confrontational, collaborative approach in working closely with governments, the business community and others” [@ChinaPhilanthropyReserchInstitute:2018a]. Over the years, they have provided technical support for the central government and local governments like the Yunnan and Zhejiang Province to build national parks and nature reserve areas and helped drafted such documents as the China Biodiversity Conservation Strategy and Action Plan (2011-2030). Using the non-confrontational working model and establishing long partnerships with the government enabled TNC to secure the consent of a national governmental agency in corresponding issue area to be their PSU, and thus successfully got registered and authorized to operate in a relatively large number of provinces. In contrast, Greenpeace states in their mission that they work as “independent campaigning organizations that use peaceful protest and creative confrontation to expose global environmental problems”. Their description of the work in China also uses such key words as “brings change”, “non-violent direct action”, and “impartiality and independence”. One project Greenpeace shares on their website was the exposure of the damages the company Sinar Mas Group's did in southwest Yunnan back in 2004. They conducted three rounds of  independent investigations and later revealed that the multinational paper industry company used reforestation as a cover while actually logging Yunnan's natural forests. Although this confrontation was not with the government, it showcases Greenpeace’s investigative, watchdog like work stance. To the authority, it would mean less predictable and manageable.

>>>>

# Civil society laws as authoritarian policy tools for shaping INGO operational space

The 2017 ONGO Law's provision of two legal pathways for registration—each with different degrees of freedom—is an example of the broader phenomenon of closing civic space. In the past two decades, authoritarian states have used administrative and legal measures to control and curtail the influence of global civil society. A rich literature in comparative politics explores how authoritarian regimes counterintuitively use democratic institutions for their own benefit and survival [@Kendall-TaylorFrantz:2014a; @GandhiPrzeworski:2007]. Authoritarian states hold elections [@LevitskyWay:2010], permit popular protests [@Weiss:2013], and allow opposition parties in the legislature and set executive term limits [@Meng:2020]. However, autocrats use these democratic-appearing institutions to limit possible threats to regime stability. Elections, protests, and apparent political reform are allowed, but only in ways that bolster the government's stability, and these institutions are kept weak and "dependent on the regime to ensure that they do not develop any real power or autonomy" [@FrantzEzrow:2011, 7].

Civil society organizations are another political institution that autocrats must balance as part of their stability-seeking calculus. Despite the fact that NGOs can act as "agents of democratization" [@ToeplerZimmerFrohlich:2020], thus posing a threat to non-democratic rule, most authoritarian states permit the growth of domestic and international civil society [@DeMattee:2019; @Heiss:2019; @Heiss:2019a]. INGOs provide needed resources and expertise in humanitarian relief, environmental policy, education, development, and autocrats can rely on these services to strengthen their regime by providing better public goods. At the same time, INGOs can pose a threat to stability by exposing human rights abuses, encouraging unwanted political reform, and engaging in political advocacy [@Heiss:2019; @ToeplerZimmerFrohlich:2020]. To capture the benefits and reduce the risk of working with global civil society, most authoritarian states regulate civil society in nonviolent ways. Instead of beating activists or arresting NGO leaders, states have relied on "administrative crackdown," [@Chaudhry:2022] or laws that create barriers to NGO advocacy, entry, and funding [@ChristensenWeinstein:2013; @ChaudhryHeiss:2022; @CarothersBrechenmacher:2014].

These laws, however, are a blunt tool that tend to target civil society as a whole—authoritarian states do not create one set of formal laws that applies only to regime-friendly INGOs and another set that applies to more risky INGOs. Instead, there is a divergence between the formal *de jure* laws and the *de facto* implementation of their accompanying regulations. In practice, NGO laws have tended to not be applied universally, and formal laws often can serve as a "weapon hanging on the wall that never fires," acting a warning to potentially uncooperative NGOs [@Kozenko:2015]. Autocrats engage with INGOs in diverging ways, providing financial support and opportunities for collaboration with those that align with government policy preferences, and using legal mechanisms to restrict or expel those that do not [@ToeplerZimmerFrohlich:2020; @Plantan:2022; @LiFarid:2023].

Laws act as policy tools that allow the government to TODO allow host governments to harness the aspects of INGO programming that align with their domestic policy preferences while limiting INGO behavior that could threaten regime stability. 
authoritarian civil society regulations act as policy tools that regimes can leverage to shift INGO behavior in their favor and how restrictive laws can reshape and tame organizational strategies [@Li:2020; @WuNie:2021; @LianMurdie:2023].
*note* changed the sub-title to incorporate the policy tool lense too.

TODO: Framework

# INGO operating space in authoritarian contexts

Anti-INGO laws thus creating different degrees of INGO "operating space," or the TODO.

We propose a framework for understanding INGO operating space

We look at operating space and some of the factors that shape it. The law is designed in a way to give the government the power to selectively grow or shrink the operating space available to INGOs, and organizations work within those bounds to maximize that space

```{r fig-theory, fig.width=6, fig.height=1.75}
#| echo: false
#| include: true
#| warning: false
#| message: false
#| fig-cap: "Diagram of our theoretical framework showing three possible determinants of INGO operational space: the government's interest in and relationship with the INGO, the contentiousness of the INGO's issue area, and the inherent geographic scope of the INGO's operations."
theory_nodes <- tribble(
  ~id, ~name, ~x, ~y, ~highlight,
  1,   "Issue contentiousness", 2, 1, FALSE,
  2,   "Operational space", 2, 0, TRUE,
  3,   "Local focus", 3, 1, FALSE,
  4,   "Government preferences", 1, 1, FALSE
)

TODO: Issue area and contentiousness and government preferences **and realignment through policy**

Finally, the natural scope of an INGO's activities can shape its allowed operational space outside of authoritarian stability seeking calculus. Many INGOs purposely keep themselves small, non-confrontational, and non-political and limit themselves to work in only a small part of their target countries, having no plans to go beyond their hyperlocal scope. This naturally limits their operational space, but the limits are self-imposed and not directly related to government regulations.

@@

# Exploring INGO operating space in China

Below, we apply this framework to China

Why China: in-depth look at the implementation of a single formal anti-INGO law and how it has reshaped the operating space available to global civil society
China provides a unique case for looking at the the exact mechanism for controlling and leveraging global civil society as an act of authoritarian stability seeking.

@@

With this framework and theoretical background, we use the case of China to explore three theoretical expectations. 
Based on extant literature and theoretical understanding, we argue that the operation space INGOs are granted is associated with the level of contentiousness their missions, i.e., the issue areas they work in. Since the level of contentiousness is also manifested by the level of difficulty they face while seeking lawful registration.

@@

TODO Andrew: Make it clear that this section is just about the data for the statsy stuff, not the case study

@@

We lack sufficient theory or data to explain why this trend goes against our expectations, but we can again make reasonable speculations. As we saw previously with INGOs and local connections, organizations appear to strategically seek out the best avenues and methods for registering [@Jia:2017a]. Organizations that participated in the initial wave of registrations in 2017 likely had closer connections with local officials and were able to speed through the bureaucratic process registration more easily. As a preliminary test of this speculation, we rerun our main model with an interaction between years since registration and local connections to see how the effect of local connections changes over time. As seen in @fig-timing-local, early registrants in 2017 without local connections are more likely to have a larger geographic reach, registering in an average of 19--23 provinces, while their locally-focused counterparts are only expected to register in an average of 7--16 provinces. Accordingly, it may be that organizations with local connections are more likely to register early because of their pre-existing bureaucratic relationships, while organizations with a larger national reach take longer to undertake the registration process and thus register later. Another possible explanation is interpreting earlier timing in registration as more eager to register as discussed in the measurement section. INGOs that self-perceive as more contentious would be more nervous about the legal consequences of not being able to register and thus be more eager to seek registration earlier [@ChinaPhilanthropyReserchInstitute:2018]. Compared with their more "rest-assured" peers, INGOs seeking to register earlier are more likely to have reduced operational space. To understand the underlying mechanism with registration timing, further research is warranted.

@@

## Contentiousness and government relationships: The cases of The Nature Conservancy and Greenpeace

@@

What factors could have contributed to the contrasting experiences of TNC and Greenpeace under the ONGO Law? After analyzing the data, the working approach and relationship with the government emerged as pivotal differences between the two INGOs, which fits existing discussions on the differential treatment of INGOs based on their relationship with the authority [TODO: add citations]. In a public interview with TNC’s RO in China about their registration experience, the chief representative stated TNC’s working approach in China has been “science-based”, using “non-confrontational, collaborative approach in working closely with governments, the business community and others” [@ChinaPhilanthropyReserchInstitute:2018a]. Over the years, they have provided technical support for the central government and local governments like the Yunnan and Zhejiang Province to build national parks and nature reserve areas and helped drafted such documents as the China Biodiversity Conservation Strategy and Action Plan (2011-2030). Using the non-confrontational working model and establishing long partnerships with the government enabled TNC to secure the consent of a national governmental agency in corresponding issue area to be their PSU, and thus successfully got registered and authorized to operate in a relatively large number of provinces. In contrast, Greenpeace states in their mission that they work as “independent campaigning organizations that use peaceful protest and creative confrontation to expose global environmental problems”. Their description of the work in China also uses such key words as “brings change”, “non-violent direct action”, and “impartiality and independence”. One project Greenpeace shares on their website was the exposure of the damages the company Sinar Mas Group's did in southwest Yunnan back in 2004. They conducted three rounds of  independent investigations and later revealed that the multinational paper industry company used reforestation as a cover while actually logging Yunnan's natural forests. Although this confrontation was not with the government, it showcases Greenpeace’s investigative, watchdog like work stance. To the authority, it would mean less predictable and manageable.
